### 2013-07-24 draft.tex
changed to elsarticle, added frontmatter, deleted big tables from the end

\documentclass[preprint, authoryear]{elsarticle}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{rotating}
\usepackage{multirow}
\usepackage{enumerate}
\usepackage{wasysym}
\usepackage{natbib}
\usepackage{lineno}

\newcommand{\deltao}{$\delta^{18}$}
\newcommand{\deltac}{$\delta^{13}$}
\newcommand{\degrees}{$^{\circ}$}

\begin{document}

\begin{frontmatter}

\title{Carbonate clumped isotope constraints on Silurian ocean temperature and \deltao O}

\author[caltech]{Renata C. Cummins\corref{cor1}}
\ead{renatacummins@gmail.com}

\author[berkley]{Seth Finnegan}

\author[washU]{David A. Fike}

\author[caltech]{John M. Eiler}

\author[caltech]{Woodward W. Fischer}

\cortext[cor1]{Corresponding author}

\address[caltech]{California Institute of Technology, Geological and Planetary Sciences, MC 100-23, Pasadena, CA 91125}

\address[berkley]{Department of Integrative Biology, University of California, 1005 Valley Life Sciences Bldg \#3140, Berkeley, CA 94720}

\address[washU]{Department of Earth and Planetary Sciences, Washington University, St. Louis, Missouri 63130}

\begin{abstract}
Much of what we know about the history of Earth's climate derives from the geologic record of carbonates. The oxygen isotopic composition (\deltao O) of carbonates has been widely used as a proxy for past ocean temperatures, but application of this proxy to deep geologic time is complicated by diagenetic alteration and uncertainties in the \deltao O of seawater in the past. Carbonate clumped isotope thermometry provides an independent estimate of the temperature of the water from which a carbonate precipitated, allowing direct calculation of the \deltao O of the water as well. The clumped isotope composition of carbonates is also highly sensitive to diagenetic alteration, enabling identification of the most pristine samples from within a sample set. These highly pristine samples ultimately provide a more reliable record of paleoclimate.

We tested the ability of carbonate clumped isotopes to constrain paleoclimate by measuring the clumped isotope composition of apparently well-preserved calcitic fossils from several Silurian-age (ca. 427 Ma) stratigraphic sections on the island of Gotland, Sweden. Variability in clumped isotope temperatures suggests differential preservation, complicating a resolved temperature history through these sections. Micro-scale studies of calcite fabric and trace metal composition using electron backscatter diffraction and electron microprobe analysis point to a subset of relatively pristine samples, which indicate that Silurian tropical oceans were likely warm ($30\pm 2$\degrees C) and similar in oxygen isotopic composition to a modern ice-free world (\deltao O$_{\text{SMOW}}$ of $-1.7\pm 0.5\permil$). This result joins the growing body of evidence that suggests the $\delta^{18}$O of Earth's surface waters has remained broadly constant through time. 
\end{abstract}

\end{frontmatter}

\linenumbers

\section{Introduction}

Despite the importance of paleoclimate and more than half a century of investigations on the topic, the temperature history of Earth's oceans remains poorly constrained. The most widely-used proxy for past ocean temperatures is the carbonate-water oxygen isotope exchange thermometer, also known as the carbonate \deltao O thermometer, which was developed by Harold Urey in the 1940s \citep{Urey1947}. This carbonate \deltao O proxy has two significant sources of complications. First, as with many paleoenvironmental proxies, it is often difficult to distinguish the effects of diagenetic alteration from the primary environmental signature. Second, the measured \deltao O of carbonates depends not only on the temperature of the seawater from which the carbonates precipitated, but also on the \deltao O of the seawater. The \deltao O of seawater can be affected by several factors, including evaporation, mixing with freshwater, and changes in the size of continental glaciers, all of which are difficult to quantify in the geologic record. The 
\deltao O of the ocean has also been hypothesized to change gradually but substantially through Earth's history \citep{Jaffres2007,Veizer1999}.

Gradual changes in the \deltao O of seawater on 100-million-year timescales seem plausible given the current knowledge of the controls on the \deltao O of seawater. The long-term average \deltao O of all the water on Earth's surface is almost entirely controlled by the balance between two exchange processes: exchange of oxygen between water and rocks during high-temperature alteration of the seafloor near mid-ocean ridges, and exchange of oxygen between water and rocks during low-temperature weathering of oceanic and continental crust. High-temperature exchange processes have small oxygen isotope fractionation factors and tend to increase the \deltao O of ocean water by equilibrating the water with the average mantle \deltao O of 5.7$\permil$ \citep{Gregory1991}. Low-temperature exchange processes have larger oxygen isotope fractionation factors and tend to decrease the \deltao O of seawater by preferentially moving $^{18}$O from water into rocks. The relative magnitudes of high- and low-temperature weathering are likely to have changed over the Phanerozoic due to uncoupled changes in the seafloor spreading rate \citep{Fornari1995,Tajika1993,Wallmann2001} and the magnitude of contintental weathering, which would lead to slow shifts in the \deltao O of the ocean. The best geologic evidence for gradual change in seawater \deltao O, however, is the subject of a long-standing controversy in geochemistry.

This controversy arises from the complications associated with the carbonate \deltao O proxy, which allow conflicting interpretations of the $\sim 8\permil$ increase in the average \deltao O of marine carbonates from the Cambrian to today. Several studies have proposed that early Paleozoic carbonates may have low \deltao O values because they are more likely to have undergone diagenetic alteration \citep{Clayton1959,Keith1964,Land1995}, which tends to decrease the \deltao O of carbonate \citep{Marshall1992}. In response to this concern, \cite{Veizer1999} measured the \deltao O of calcite fossils that were screened for diagenetic alteration using trace metal compositions and micro-scale studies of calcite fabrics. These screened fossils also record increasing \deltao O through the Phanerozoic, arguing against diagenetic alteration as a cause of the low \deltao O values in the early Paleozoic. An alternative hypothesis holds that oceans could have been much warmer in the early Paleozoic \citep{Karhu1986,Knauth1976}. Assuming that the \deltao O of the oceans was the same as today, the early Paleozoic oceans must have been $\sim 56$\degrees C to produce carbonates with a \deltao O$_{\text{PDB}}$ of $-8\permil$. These warm temperatures seem incompatible with evidence for a large glaciation in the Late Ordovician \citep{Veizer1986}, leading some studies to suggest a third hypothesis, that the \deltao O$_{\text{SMOW}}$ of seawater has increased from $-8\permil$ in the Cambrian to its present value of $\sim 0\permil$ \citep{Jaffres2007,Veizer1999}.

Carbonate clumped isotope thermometry provides a unique insight into the controversy described above because it resolves the two main complications of the carbonate \deltao O proxy. The clumped isotope composition of a carbonate records a temperature that is independent of the \deltao O of water and highly sensitive to diagenetic alteration. The extent to which $^{13}$C and $^{18}$O occur together, or ``clump'' within the same carbonate ion is a function of the temperature at which the carbonate precipitates. For a shelly marine organism that lives in shallow water, the degree of clumping in the shell calcite will reflect the sea surface temperature \citep{Ghosh2006}. Measurement of the \deltao O of the shell calcite in addition to the temperature allows calculation of the \deltao O of the seawater, using the experimentally-determined temperature dependence of the oxygen isotope fractionation between water and calcite \citep{Kim1997}. If the calcite shell is then buried and fossilized, it is often altered in the presence of high temperatures. This alteration can take the form of recrystallization or thermally-activated re-ordering \citep{Eiler2011}. In either case, the altered material will take on a new clumped-isotope temperature that can be much higher than the temperature of the pristine fossil calcite. If the altered material is mixed with pristine fossil material during sample processing, the clumped isotope temperature of the sample will be significantly higher than the temperature of the ocean water where the organism lived. We demonstrate here that carbonate clumped isotopes can be a more sensitive indicator of diagenetic alteration than the \deltao O or the trace metal content of a carbonate.

In addition to being highly sensitive to diagenetic alteration, carbonate clumped isotopes are also diagnostic of diagenetic regimes, ultimately allowing identification of the most pristine endmember of a sample set. Clumped isotopes can distinguish between diagenesis in the presence of low and high water-to-rock ratios (rock-buffered and water-buffered diagenesis, respectively) and diagenesis in the presence of meteoric water. As carbonates are buried deeper and the temperature increases, the equilibrium oxygen isotope fractionation between water and carbonate grows smaller, preferentially driving $^{18}$O into the water. During rock-buffered diagenesis, this causes the \deltao O of the water to increase, while the \deltao O of the altered carbonate is not affected. In a water-buffered diagenetic regime, the \deltao O of the water does not change, so increasing temperature instead causes the \deltao O of the altered carbonate to decrease. In contrast, diagenesis in the presence of meteoric water usually occurs when the sample is near the Earth's surface, so the clumped isotope temperature of the altered carbonate is cool, but the \deltao O of the carbonate and the water is diagnostically low. The ability of clumped isotopes to constrain the conditions to which fossils have been exposed after burial allows identification of the best-preserved samples from within a set of samples that traditionally would all be considered pristine \citep{Eiler2011}. We demonstrate here that this distinction has important implications for the estimation of the temperature and \deltao O of oceans in the past.

The present study investigates the implications of diagenetic alteration for the resolution of small temperature changes recorded in relatively well-preserved Paleozoic rocks. This study also provides a test of established proxies for diagenetic alteration as filters for clumped isotope data. Finally, our data add to the clumped isotope record of past ocean temperatures and \deltao O, including samples from a different continent and a different paleocean basin than previous clumped isotope studies.

\section{Geological setting and materials}

Samples were collected from a section exposed on the island of Gotland, Sweden, that spans the latest Llandovery through the Wenlock in the Early Silurian (Figure \ref{map}). Ages of the formations and correlations between outcrops are constrained by conodont and graptolite biostratigraphy \citep{Jeppsson2006}. The section records shelf, carbonate platform and backreef lagoon sediments deposited on the margin of the Baltic basin, which was within 30$^{\circ}$ of the equator during the Silurian \citep{Torsvik1992}.

The oldest exposed strata on Gotland are the Lower Visby Beds, which consist of marls interbedded with fine-grained limestones and were likely deposited below storm wave base (Figure \ref{column}). Fossils are present but not abundant \citep{Calner2004a,Samtleben1996}. The contact between Lower and Upper Visby beds is marked by a bed with a dense population of \textit{Phaulactis}, a species of solitary rugose coral \citep{Jeppsson1997,Jeppsson2006,Munnecke2003,Samtleben1996}. This contact is just above the Llandovery-Wenlock boundary, based on conodont stratigraphy \citep{Aldridge1993,Jeppsson1983,Mabillard1985}. During sampling, we defined this contact as a stratigraphic height of zero (Figure \ref{column}). In addition to the bed of abundant \textit{Phaulactis} corals, the contact between Lower and Upper Visby Beds is also marked by a facies change that represents a maximum flooding surface \citep{Calner2004b}. Carbonate content increases in the Upper Visby Beds, with a decreasing frequency of marl beds and the appearance of detritic limestones and reef mounds. Ripple marks indicate deposition above storm wave base. Fossils are abundant in the Upper Visby Beds \citep{Calner2004a,Samtleben1996}.

Above the Upper Visby Beds, the H\"{o}gklint Formation consists of algal and crinoidal limestones with some reef mounds, deposited above storm wave base \citep{Riding1991,Samtleben1996,Watts2000}. The contact between the H\"{o}gklint Formation and the overlying Tofta Formation represents a hypothesized sequence boundary \citep{Calner2004b}. The Tofta Formation consists of oncoid-rich bedded limestone, and likely represents a shallow restricted setting, such as a backreef lagoon \citep{Riding1991,Samtleben1996}. The overlying Hangvar Formation consists of interbedded marls and limestones with some reef mounds, indicating deposition in a less restricted, deeper-water setting. The youngest strata sampled for this study were from the Slite Formation, which transitions from limestone-rich to marl-rich interbedded limestones and marls, likely representing deposition in deeper water. Fossils are present throughout the section \citep{Calner2004a}.

The Silurian section on Gotland is an attractive location to sample for a clumped isotope study for several reasons. First, this section has long been recognized for its excellent preservation. The strata on Gotland are very nearly flat-lying, with a dip of less than 1\degrees, and display only minimal evidence of tectonic disturbance \citep{Calner2004a,Jeppsson1983}. Conodont fossils from this section are pale yellow in color, giving them a conodont alteration index (CAI) of one, which corresponds to a maximum temperature of $\sim$100\degrees C \citep{Epstein1976,Jeppsson1983,Wenzel2000}.

The Silurian section on Gotland is also attractive due to the relative abundance of fossils, especially brachiopods, throughout the section. By sampling fossil material for clumped isotope measurements, we target calcite that was first precipitated directly from seawater. All brachiopod samples were taken from the secondary interior fibrous layer of the brachiopod shell, which is thought to precipitate in equilibrium with seawater and to be relatively resistant to diagenetic alteration \citep{Azmy1998,Samtleben2001}. The fibrous calcite fabric characteristic of the secondary layer has been well-characterized in modern and fossil brachiopods \citep{PerezHuerta2007,Samtleben2001,Schmahl2004}, providing an expectation for pristine samples in our dataset. The trace metal composition of brachiopods has also been extensively studied as an indicator of diagenetic alteration \citep{Azmy1998,Brand2003,Brand2012,Mii1994,Grossman1996,Shields2003}. The other fossils present in this section, including rugose corals, crinoids, bryozoans, and gastropods, are used less often for paleoenvironmental studies than brachiopods, but have been used by other clumped isotope studies of paleotemperature \citep{Came2007,Dennis2013,Finnegan2011}. Measurements of several fossil taxa allow comparison of the preservation quality of different calcite materials.

A third attractive feature of the Silurian section on Gotland is the coupled positive excursion in \deltac C and \deltao O across the boundary between the Lower and Upper Visby Beds (Figure \ref{column}). This excursion is called the Ireviken event, and is recognized globally \citep{Munnecke2003}. In Gotland, the 3$\permil$ excursion in \deltac C and 0.6$\permil$ excursion in \deltao O reach a maximum rate of change across the contact between the Lower and Upper Visby Beds \citep{Munnecke2003}. The excursions coincide with a shift to deposition in shallower water, but it is difficult to distinguish the relative contributions of progradation and eustatic sea level fall to this facies change \citep{Calner2004a}. The excursions are also broadly coincident with minor extinctions in several benthic and pelagic marine taxa \citep{Munnecke2003}.

The coincident isotopic excursions and stratigraphic change suggest that the Ireviken represents a climate phenomenon, the nature of which is still debated. The similarity between the isotopic signatures of the Ireviken event and known glaciations, such as the Late Ordovician glaciation, have led several studies to hypothesize a glacial cause for the Ireviken event \citep{Azmy1998,Kaljo2003,Brand2006,Calner2008}. This hypothesis is supported by the widespread presence of glacial tillites in Brazil, dated to the latest Llandovery or earliest Wenlock \citep{Grahn1992}. Many of these observations could also be explained by a shift from a humid period to an arid period in the tropics \citep{Bickert1997,Munnecke2003,Samtleben1996}. According to this model, during the humid period recorded by the Lower Visby Beds, increased precipitation would lower the \deltao O of surface waters and also stimulate the upwelling of deep water with low \deltac C onto the shelves. Increased precipitation would also cause intense weathering, which explains the dominance of marls over carbonate reefs during this time. Then during the arid period beginning in the Upper Visby Beds, evaporation would increase the \deltao O of surface waters, and downwelling of saline water would bring only surface water with high \deltac C onto the shelves. Low terrigenous influxes would allow the growth of carbonate reefs, observed in the H\"{o}gklint Formation.

In the present study, we test the ability of clumped isotopes to resolve the changes in ocean temperature and seawater \deltao O that contributed to the 0.6$\permil$ shift in the \deltao O of carbonates through the Ireviken event. From first principles, it will be difficult for the carbonate clumped isotope thermometer to resolve these changes, because a shift of 0.6$\permil$ in \deltao O corresponds to a change in temperature of approximately 2\degrees C, which is near the limit of resolution of carbonate clumped isotopes.

\section{Methods}

\subsection{Collection and preparation of samples}

Samples were collected from the island of Gotland, Sweden during the summer of 2011. Each sample was labeled with a location, as shown in Figure \ref{map}, and a stratigraphic height. For locations where the \textit{Phaulactis} layer is exposed, stratigraphic height was expressed relative to the boundary between the Lower and Upper Visby Beds (Figure \ref{column}). Other locations were correlated on the basis of prior sequence-, bio- and chemostratigraphic work \citep{Calner2004a}.

Most of the fossils were easily exposed by breaking the surrounding layers of poorly-lithified marl. For fossils in more strongly lithified beds, fossil material was accessed by cutting away the rock with a saw or microrotary drill. After fossils were separated from the rock, they were sonicated for two hours to remove residual sand, silt and mud from the surface of the fossils. Brachiopods, bryozoans, crinoids, gastropods and ostracods were sampled by flaking the surface of the skeleton or shell using a dental chisel. Every flake was visually inspected under a microscope at circa $25\times$ power. Flakes with visible pyrite grains or rough-textured recrystallized regions were rejected and not included in the material for analysis. For brachiopods, only the flakes preserving the characteristic fabric of the fibrous secondary layer were chosen for analysis (see Figure \ref{flakes}). Flakes were then sonicated briefly, rinsed, dried in a 50\degrees C oven, and powdered using a mortar and pestle. Only the apical ends of rugose corals were sampled, after removal of any sediment from the surface using an abrasive bit. The apical end was then cut off using a diamond microsaw, and powdered using a mortar and pestle. The apical ends of the corals were targeted because the coral septa are closely spaced together there, leaving less room for cement to precipitate between the septa. However, this sampling technique is still not as selective as the technique used for brachiopods, and may include more cement and sediment along with skeletal calcite. Finally, to assess diagenetic alteration, micrite and cement samples were drilled from cut surfaces and then powdered using a mortar and pestle.

While all fossils were inspected under a microscope prior to sampling, we were not able to inspect all fossils in thin section, because carbonate clumped isotope measurements require large amounts of material, some of which is lost during preparation of thin sections. However, 30 $\mu$m-thick uncovered, polished thin sections were obtained from cuts through four representative fossils, two brachiopods and two rugose corals. These thin sections were inspected first in transmitted and reflected light to characterize calcite textures, and were then examined using several different spectroscopic techniques, described below.

\subsection{Electron microscopy}

Electron backscatter diffraction (EBSD) provides an independent measure of recrystallization by mapping the orientation of calcite crystals in a thin section. This allows more detailed characterization of calcite fabrics than is possible using optical microscopy, and in the case of brachiopods, comparison to pristine modern examples \citep{PerezHuerta2007}. Two thin sections, one rugose coral and one brachiopod, were analyzed using a Zeiss 1550VP Field Emission Scanning Electron Microscrope (SEM) under variable pressure vacuum, with an electron high tension of 20 kV. The stage was tilted 70$^{\circ}$. The AZtec program was used in $4\times4$ binning mode, with frame averaging over one frame and a step size of 2 $\mu$m for the rugose coral and 1.2 $\mu$m for the brachiopod. The electron beam interacts with the crystal and is diffracted in a characteristic pattern for calcite. The diffracted beam then interacts with the EBSD detector, depending on the orientation of the calcite crystal. The data were analyzed using software from Oxford instruments.

Flakes of the secondary fibrous layer from over ten different brachiopods were also imaged to further inspect the preservation of original fibrous calcite texture. The rinsed and dried flakes were prepared for imaging by coating them with a layer of carbon 12 nm thick. The coated flakes were imaged using the secondary electron detector in the same SEM as above, under high vacuum. The working distance was $\sim 6$ mm and the electron high tension was 10 kV.

\subsection{Wavelength-dispersive spectroscopy}

The use of wavelength-dispersive spectroscopy in combination with optical microscopy and EBSD allows characterization of the trace metal contents of specific calcite textures. By comparing the trace metal concentrations of primary textures to those of diagenetic textures, we can identify the trace metal signature of diagenetic alteration in our sample set. The same thin sections as above were coated with a layer of carbon 15 nm thick, and then analyzed with wavelength-dispersive spectroscopy using the JEOL JXA-8200 Electron Probe Micro-analyzer (E-probe). Using the JEOL map analysis program, the beam was calibrated on crystals of dolomite, siderite, strontianite and rhodochrosite. Qualitative maps of Ca, Mg, Fe, Sr and Mn concentration were made using a current of 200 nAmps, voltage of 15 kV, and a counting time of 80 milliseconds per pixel. Quantitative spot measurements were made using the JEOL quantitative analysis program.

\subsection{Bulk trace metal analysis}

Although we attempted to sample only well-preserved skeletal material for the carbonate clumped isotope measurements, the large mass of material required makes it very difficult to avoid including small amounts of other phases. The trace metal concentrations of bulk samples can be used as an indicator of the amount of sediment and recrystallized material included in the sample \citep{Azmy1998,Brand2012,Came2007,Finnegan2011}. Diagenetic recrystallization of calcite is thought to enrich iron (Fe) and manganese (Mn) and deplete strontium (Sr) \citep{Brand1980,Shields2003}, so we measured the bulk concentrations of these three elements in a large and representative subset of our samples. Between 1 and 5 milligrams of each powdered sample were dissolved in a known volume of 4\% nitric acid (HNO$_3$). Standards were prepared from standard solutions of Sr, Mn and Fe with the same concentration of Ca and HNO$_3$ as the sample solutions to minimize matrix effects. Solutions were sampled using a Cetac ASX 260 autosampler, aspirated into the Ar plasma using a peristaltic pump, and then analyzed on a Thermo Scientific iCAP 6000 series inductively-coupled plasma optical emission spectrometer (ICP-OES). Between each set of ten unknown samples, three standard solutions (10 ppb, 100 ppb, and 1 ppm) were measured. Reproducibility was on average 3\% for Fe, 2\% for Mn and 3\% for Sr.

\subsection{Clumped isotope analysis and standardization}

Each sample was measured between one and three times. Approximately 10 milligrams of powdered sample were dissolved in phosphoric acid at 90\degrees C, which produces CO$_2$ with a known oxygen isotopic offset from the sample carbonate. The subsequent cleaning of the CO$_2$ followed procedures outlined by \cite{Ghosh2006}. The cleaned CO$_2$ was analyzed on a Finnigan MAT253 gas source mass spectrometer configured to collect masses 44 through 49. The sample gas was measured in eight separate acquisitions, each of which was bracketed by a measurement of reference CO$_2$ with known \deltac C and \deltao O. The \deltac C, \deltao O, $\delta_{45}$, $\delta_{46}$, $\delta_{47}$, and $\delta_{48}$ were calculated for each sample relative to the reference CO$_2$. Acquisition-to-acquisition standard deviations were on average 0.03$\permil$ for $\delta_{47}$, 0.04$\permil$ for \deltao O, and 0.02$\permil$ for \deltac C. Raw measured values for $\Delta_{47}$ and $\Delta_{48}$ of each sample, defined as $\Delta_{47-[\text{SGvsWG}]}$ and $\Delta_{48-[\text{SGvsWG}]}$, were also calculated relative to the reference CO$_2$ following equations 3 and 4 in \cite{Huntington2009}.

All $\Delta_{47}$ values are projected into the absolute reference frame \citep{Dennis2011}. When possible, raw $\Delta_{47-[\text{SGvsWG}]}$ values are converted to the absolute reference frame using an empirical transfer function (ETF) constructed from the heated and equilibrated gases measured during the same week as the samples. Heated gases are aliquots of CO$_2$ that were sealed in quartz tubes, heated at 1000\degrees C for three hours, and then cooled rapidly to room temperature. Equilibrated gases are aliquots of CO$_2$ that were equilibrated with water at 25\degrees C for at least 24 hours. Heated gases were made from two separate reservoirs of CO$_2$ with different \deltao O compositions, and equilibrated gases were equilibrated with either of two separate waters, also differing in their \deltao O compositions. Heated and equilibrated gases were cleaned following the same procedure as the samples. The ETF was constructed and then used to convert raw $\Delta_{47-[\text{SGvsWG}]}$ values to the absolute reference frame following \cite{Dennis2011}. Finally, the resulting $\Delta_{47-\text{RF}}$ was corrected for the acid extraction at 90\degrees C by adding 0.092$\permil$ \citep{Henkes2013}.

Some samples were measured during weeks when no equilibrated gases were measured, so an ETF could not be constructed for those weeks. These samples were projected into the absolute reference frame using a secondary transfer function (STF). Accepted values of two California Institute of Technology intralab standards were determined in the absolute reference frame from measurements of these standards during weeks when an ETF could be constructed from heated and equilibrated gas measurements. To create the STF, the accepted values for the two standards, as well as the accepted value of 0.0266$\permil$ for heated gases, were plotted against their $\Delta_{47}$ as determined in the \cite{Ghosh2006} reference frame. $\Delta_{47}$ values of the samples in the \cite{Ghosh2006} reference frame were corrected using this STF following equation C1 in \cite{Dennis2011}. A separate STF was constructed for each week of measurements during which no equilibrated gases were measured.

Although all samples were cleaned prior to introduction into the mass spectrometer, the possibility for contamination from hydrocarbons and chlorocarbons still exists. To screen for contamination, the $\delta_{48}$ and $\Delta_{48}$ of each sample were compared to a line of best fit through a plot of $\delta_{48}$ versus $\Delta_{48}$ of the heated gases, all in the \cite{Ghosh2006} reference frame. Samples whose $\Delta_{48}$ fell more than 1$\permil$ off the line were discarded due to likely contamination.

$\Delta_{47-\text{RF-AC}}$ values were converted to temperature using the \cite{Ghosh2006} calibration line, which is projected into the absolute reference frame in equation 9 in \cite{Dennis2011}. The \deltao O of water was calculated from the temperature and the measured \deltao O of the sample using the temperature-dependent fractionation, $\alpha$, of oxygen isotopes between water and calcite, $1000\ln\alpha=18.03\left(10^3\text{T}^{-1}\right)-32.42$, where T is temperature in degrees Kelvin \citep{Kim1997}.

\section{Results and discussion}

\subsection{Stratigraphic trends}

The \deltac C values of fossil material, micrite and cement are highly consistent within beds (Figure \ref{section}). We resolve a 3$\permil$ positive excursion in \deltac C, beginning just below the boundary between the Lower and Upper Visby Beds, and continuing for at least 50 meters. The sample set does not capture the relaxation of this excursion, but samples from the Slite Formation, beginning at 76 meters, record a return to lower \deltac C$_{\text{PDB}}$ values of approximately -0.5$\permil$. The \deltac C trend through the section is comparable to the brachiopod \deltac C measurements from \cite{Bickert1997} and \cite{Munnecke2003}.

The \deltao O values of fossil material are consistent within beds, but micrite and cements tend to have lower \deltao O values than fossil material, often several permil lower than fossils in the same bed (Figure \ref{section}). We are not able to resolve an excursion in \deltao O, although this may simply be because our dataset contains fewer samples than that of \cite{Munnecke2003}. The average \deltao O$_{\text{PDB}}$ of the fossils remains fairly constant at -5.5$\permil$ through the Lower and Upper Visby Beds, making our Lower Visby samples consistent with those of \cite{Munnecke2003} but our Upper Visby samples more depleted in $^{18}$O. Higher in the section, the \deltao O of the fossils in this study appear consistent with values reported by \cite{Bickert1997}.

The clumped isotope temperature of samples is highly variable within beds, often covering a range of 10\degrees C within a single bed (Figure \ref{section}). The lowest temperature sample, a brachiopod, and the highest temperature sample, a rugose coral, are both from the Slite Formation, deposited well after the hypothesized climate phenomenon associated with the isotopic excursions in the Lower and Upper Visby Beds. Due to the scatter in temperature, we are not able to resolve a change in temperature with time.

The large scatter in temperatures suggests variable preservation quality among the samples. Even brachiopods, which are commonly thought to be resistant to diagenetic alteration, have a range in temperature of more than 15\degrees C. All of the fossil samples have fairly high values of \deltao O, making alteration by meteoric water unlikely. This leads to a hypothesis that the lowest temperature samples are the most pristine, while the higher temperature samples likely represent diagenetic alteration during burial at high temperature. We tested this hypothesis using independent methods to assess the preservation quality of representative samples.

\subsection{Independent assessments of preservation quality}

Bulk trace metal measurements, which have been widely used as indicators of diagenetic alteration, suggest that all the brachiopods in this study are pristine. The bulk Mn and Sr concentrations of all brachiopods fall within previously defined limits for the Mn and Sr concentrations of pristine brachiopods (Figure \ref{Brand}). However, micro-scale studies of calcite fabrics, as well as clumped isotope measurements, demonstrate that there is a range in preservation quality among the brachiopods that cannot be identified on the basis of trace metal measurements alone. We first explore possible reasons for the insensitivity of trace metal concentrations to preservation quality. Then we consider EBSD and SEM evidence for the inclusion of diagenetic material in the typical fossil sample. Finally, we explore the use of clumped isotopes to distinguish the best-preserved samples from among those that would ordinarily all be identified as pristine based on their trace metal contents.

There are several possible explanations for the relative insensitivity of trace metal concentrations to diagenetic alteration. Our sample set may not include a sufficient number of highly altered samples to capture the full range in severity of alteration. This hypothesis is supported by a conspicuous lack of statistically robust correlations between trace metal composition and temperature (Figure \ref{metals}). We tested brachiopods and rugose corals separately and combined with micrites and cements for the robustness of a linear correlation between trace metal concentrations and temperature, as well as the log of trace metal concentrations and temperature. We performed a hypothesis test on each pair of variables to test the probability (p) that the observed correlation is due to random variation when the true correlation is zero. Only one pair of variables, the Sr concentration and temperature of rugose corals, has a p value of less than 0.05, meaning we can be 95\% confident that the correlation is real. While these tests suggest that the scatter in trace metal concentrations is a large cause of their insensitivity to diagenetic alteration, there is also the possibility that early void-filling cements, which would have cool temperatures but altered trace metal concentrations, interfere with the signal.

E-probe analysis suggests a more compelling reason for the insensitivity of trace metal concentrations: diagenetic recrystallization may not actually have any effect on Fe and Mn concentrations. A recrystallized region of a rugose coral septum is clearly visible as a dark area in reflected light (Figure \ref{EBSD_coral}A) and an area of aligned calcite crystals in an electron backscatter diffraction (EBSD) image (Figure \ref{EBSD_coral}B). However, the Fe and Mn concentrations of the recrystallized region are indistinguishable from those of the pristine coral skeleton (Figure \ref{EBSD_coral}C \& D). This is only one example, but it suggests that recrystallization may not affect trace metal concentrations, and instead, variations in trace metal concentrations may be dominated by contamination from cements and sediments. If recrystallization is an important contributor to the elevated temperatures of fossil material, trace metals may not be expected to be reliable indicators of diagenetic alteration.

We therefore turn to micro-scale studies of calcite textures as indicators of diagenetic alteration. Reflected-light and EBSD images of rugose corals reveal substantial amounts of late-phase diagenetic calcite, including cements and recrystallized regions (Figure \ref{EBSD_coral}A \& B). As discussed above, recrystallized regions are visible as dark colors in reflected light and aligned calcite crystals in EBSD images. Cements in Figure \ref{EBSD_coral}A \& B are characterized by large, subequant, interlocking calcite crystals. The fraction of late-phase calcite in the rugose coral samples may have been lower than suggested by the thin section in Figure \ref{EBSD_coral}, because samples were taken from the apical end of the corals where the void space between the septa was smaller, but sample preparation probably included some late-phase calcite along with pristine coral skeletal material.

In contrast to rugose corals, the EBSD image of a low-temperature brachiopod thin section demonstrates excellent preservation of original fibrous calcite textures (Figure \ref{EBSD_brach}A). Within the brachiopod shell, calcite crystals are elongated approximately parallel to the surface of the shell, as expected for the fibrous secondary layer. This preferential orientation is represented by the vertical band in the stereographic projection of crystallographic planes (Figure \ref{EBSD_brach}B). No regions of recrystallization are visible within the shell. There are large interlocking cement crystals in the upper left hand corner of the EBSD image, but these formed in the void space between the two valves, and would not have been included in the sample for clumped isotope analysis. Based on visual inspection of the EBSD image of this brachiopod's secondary fibrous layer, we suggest that this layer contains a maximum of 5\% recrystallized material.

Scanning electron microscope (SEM) images of low-temperature brachiopod flakes also demonstrate excellent preservation of original fibrous calcite textures (Figure \ref{flakes}A, B \& C). The original fibrous texture clearly contrasts with the rough recrystallized texture covering part of a flake that would not have been chosen for analysis (Figure \ref{flakes}D). The difference between these two textures is visible under the microscope used for picking brachiopod flakes, so flakes with large recrystallized regions were easily identified and excluded from the sample. However, small recrystallized regions, especially if they occurred between layers of fibrous texture, may not have been identified during picking. Therefore, although the representative flakes chosen for inspection with the SEM appear pristine, this does not guarantee that all recrystallized material was excluded from the samples.

The micro-scale studies of calcite textures in Figures \ref{EBSD_coral}--\ref{flakes} reveal relatively well-preserved samples, but do not rule out the inclusion of small amounts of late-phase calcite. Variable fractions of late-phase calcite likely contribute to the range in temperature among fossil samples. We only performed SEM and EBSD analysis on a small subset of samples, so we will not attempt to use this evidence to compare the relative preservation quality of low and high-temperature samples. An important conclusion that can be drawn from our micro-scale studies is that the low temperature brachiopods imaged in Figures \ref{EBSD_brach} and \ref{flakes} probably contain only trace amounts of late-phase calcite, likely less than 5\% of the sample. We therefore look to these samples as the most reliable records of ocean temperature and \deltao O during the Silurian.

\subsection{Constraining the temperature and $\mathbf{\delta^{18}}$O of Silurian oceans}

To estimate the temperature and \deltao O of Silurian oceans, we first need to confirm the hypothesis that the lowest-temperature samples are the most pristine and therefore preserve the most reliable record of Silurian ocean conditions. This hypothesis can be tested through constraints on the diagenetic regime to which the samples were subjected. As outlined above, different diagenetic regimes have characteristic effects on the relationship between temperature and the \deltao O of interstitial water. When the clumped isotope temperatures of the samples in this study are plotted versus the \deltao O of water calculated for each sample, there is a clear positive relationship (Figure \ref{ellipses}). This relationship suggests that interstitial water grew increasingly enriched in $^{18}$O as the temperature increased, which is diagnostic of rock-buffered diagenesis. Importantly, during rock-buffered diagenesis, more altered samples have higher temperatures, which is consistent with our hypothesis that the lowest temperature samples are the most pristine.

Solid-state diffusion is commonly considered as a possible mechanism for the increase of clumped isotope temperatures during diagenesis, but we rule it out as a major contributor to our results. During solid-state diffusion, isotopes of carbon and oxygen diffuse through the calcite mineral lattice, which decreases the degree of clumping and increases the interpreted temperature of the calcite. This process does not affect the bulk isotopic composition of the calcite, and it occurs without dissolution and reprecipitation, which means it is also not expected to change the calcite fabric or trace metal content \citep{Eiler2011}. However, solid-state diffusion is expected to affect all calcite approximately equally, which means that this process cannot explain the observed variability in temperature among the samples. Additionally, solid-state diffusion is thought to occur very slowly at temperatures below 100\degrees C (D. Stolper personal communication 2013), which is the maximum temperature experienced by Gotland strata, as constrained by the conodont alteration index \citep{Jeppsson1983,Wenzel2000}. Solid-state diffusion is not consistent with our observations, leaving rock-buffered diagenesis as the most likely characterization of the diagenetic regime.

The hypothesis that the samples underwent rock-buffered diagenesis is further supported by comparisons of the clumped isotope temperature of different calcite textures. We measured the clumped isotope composition of the micrite adjecent to four brachiopods, and in all four cases, the temperature of the micrite was resolvably higher than the temperature of the brachiopod (Figure \ref{comparisons}). Micrite is expected to be less pristine than the brachiopod samples because micrite is fine-grained, not protected by a layer of brachiopod shell, and was not picked under a microscope prior to analysis. Because the less pristine material, micrite, was associated with higher temperatures, these comparisons are consistent with the hypothesis that the lowest temperature samples are the most pristine.

The lowest-temperature, most pristine sample in this study, a brachiopod, indicates that tropical Silurian oceans had a temperature of $30\pm 2$\degrees C and a \deltao O$_{\text{SMOW}}$ of $-1.7\pm 0.5\permil$ (Figure \ref{ellipses}), where the error estimates represent the standard error of the mean over eight acquisitions. The two next-most pristine samples, also brachiopods, have temperatures of $33\pm 2$\degrees C and $35\pm 2$\degrees C, which correspond to seawater \deltao O$_{\text{SMOW}}$ values of $-1.5\pm 0.3\permil$ and $-1.9\pm 0.4\permil$, respectively. One caveat of these values is that they are each based on only one measurement due to sample-size limitations. The lowest-temperature brachiopod was picked a second time, but to collect enough material for analysis, some less pristine flakes had to be included, which may explain the $41\pm 4$\degrees C temperature of this re-sample (G13 10.5 to 10.8 B1 leftovers). However, we argue that the differences between the temperatures and \deltao O values implied by the three most pristine samples are small enough that they essentially represent the same warm, tropical Silurian ocean water.

A second caveat of our estimates of the temperature and \deltao O of Silurian tropical oceans is that they are technically upper estimates. We are confident that our sample suite was subjected to a diagenetic regime that increased the temperature and \deltao O of water recorded by the samples. We are not able to exclude the possibility that diagenesis affected all the samples in our study, even the most pristine samples. However, micro-scale studies of the calcite fabric of the lowest-temperature brachiopods indicate that late-phase calcite is only a trace constituent of these samples, comprising a maximum of 5\% of the material. Assuming linear mixing, which is not technically correct but close enough for our purposes, if 5\% of the sample material was recrystallized at 100\degrees C and the rest of the sample material retained an original temperature of 30\degrees C, the measured temperature of the sample would be $\sim34$\degrees C. Therefore, the constraints provided by the EBSD mapping of the brachiopod shell calcite demonstrate that our estimate of the temperature of the tropical Silurian ocean is likely reliable to within 4\degrees C.

Our clumped isotope measurements indicate that Silurian tropical oceans were similar to modern tropical oceans in their temperature and \deltao O. Modern measurements of sea surface temperature in the Solomon Sea, east of Papua New Guinea, are 30\degrees C \citep{Reynolds1994}, matching our estimate of the temperature at a similar latitude during the Silurian. The average \deltao O$_{\text{SMOW}}$ of the modern ocean is approximately $-0.3\permil$ \citep{Shackleton1974}, but without ice, the average \deltao O$_{\text{SMOW}}$ would decrease to $-1.4\permil$ \citep{Lhomme2005}. The volume of continental glaciers is poorly constrained for the Early Silurian, but regardless, the \deltao O of Silurian oceans was likely within a few permil of the modern value.

The results of this study are also consistent with those of earlier clumped isotope studies. \cite{Came2007} and \cite{Finnegan2011} both measured the clumped isotope compositions of brachiopods from the Early Silurian on Anticosti Island, Canada. These two studies report temperatures of 30 to 35\degrees C and seawater \deltao O$_{\text{SMOW}}$ values of around $-1\permil$, which are similar to the results of this study. Importantly, Anticosti Island recorded deposition on the eastern margin of the Panthalassic Ocean during the Silurian, while Gotland recorded deposition in the Rheic Ocean, which was separated from the Panthalassic Ocean by a mountain range. The strong similarity in temperatures and \deltao O between the two studies indicates that both studies likely reflect water that is representative of the global ocean, rather than an isolated inland sea.

Our results, along with those of previous clumped isotope studies \citep{Came2007,Dennis2013,Finnegan2011}, provide consistent explanations for the low \deltao O values of early Paleozoic carbonates (Figure \ref{Veizer}). Carbonate clumped isotopes constrain the contribution of a change in seawater \deltao O to only a few permil of the observed trend, ruling out a value of $-8\permil$ for the \deltao O$_{\text{SMOW}}$ of early Paleozoic oceans \citep{Grossman2012}. Our results also serve as a reminder that many of the fossils preserved in the geologic record lived in tropical oceans. The \deltao O of carbonate in equilibrium with tropical ocean water should not be compared to the \deltao O of carbonate in equilibrium with water of average modern sea surface temperature, because tropical oceans are much warmer than the average sea surface temperature. Clumped isotopes additionally demonstrate that extremely hot ($>50$\degrees C) ocean temperatures cannot be the cause of the depleted \deltao O values of early Paleozoic carbonates. Finally, clumped isotopes, as a highly sensitive indicator of diagenetic alteration, demonstrate that even among samples with original calcite fabric and pristine trace metal signatures, there is a large range in preservation quality. This means that the average fossil calcite in almost any sample suite is likely altered from its primary composition, and that diagenetic alteration is a key contributor to the low \deltao O values of early Paleozoic carbonates.

Due in large part to diagenetic alteration, our results are unable to resolve a change in temperature or the \deltao O of the oceans through the Ireviken event. The consistent scatter in fossil temperatures throughout the section likely rules out any large changes in temperature or \deltao O of the oceans, but this scatter prevents resolution of any smaller changes. Temperatures of 30\degrees C in tropical oceans are similar to those of modern tropical oceans, and therefore are not incompatible with the presence of polar ice sheets, as suggested by Early Silurian age tillites in Brazil \citep{Grahn1992}.

The results from this study join those of previous clumped isotope studies in implying that the \deltao O$_{\text{SMOW}}$ of seawater has been fairly constant throughout the Phanerozoic, at around $-1\permil$. Several studies have suggested that high-temperature alteration of the seafloor acts as a buffer for the \deltao O of seawater \citep{Muehlenbachs1976,Gregory1981,Gregory1991,Muehlenbachs1998}, but a more recent model implied that this buffer may not be strong enough to overcome slow changes in the \deltao O of seawater \citep{Jaffres2007}. If this buffer cannot explain the relative constancy of seawater \deltao O values through time, then the two competing controls on seawater \deltao O, alteration of rocks at high and low temperatures, must have remained in balance over the last 500 million years. This suggests that the magnitudes of high-temperature hydrothermal alteration and low-temperature weathering have not changed substantially, or that any changes in the magnitude of one process are tied through an as-yet unidentified feedback loop to balancing changes in the other.

\section{Conclusions}

We used clumped isotopes to constrain the temperature and \deltao O of Silurian oceans because they provide a temperature that is independent of the \deltao O of the water from which the carbonate precipitated. Samples were collected from relatively well-preserved sections on the island of Gotland, Sweden. By sampling well-preserved, fossil-rich rocks that record a climate phenomenon thought to be associated with changing seawater temperatures, we hoped to test the limit of resolution of clumped isotopes. We were not able to resolve a change in temperature through the section, due to the variable preservation quality of the samples. Trace metal analysis was not able to detect variations in preservation among brachiopods, despite a range in brachiopod temperatures, suggesting that clumped isotopes are a more sensitive indicator of diagenetic alteration than trace metal compositions. Micro-scale analysis of calcite fabrics, comparisons between brachiopods and micrite, and the positive relationship between temperature and \deltao O of water calculated from the samples all support the hypothesis that the lowest-temperature samples are the most pristine. Based on the most pristine sample in our dataset, we estimate that tropical Silurian oceans had a temperature of  $30\pm 2$\degrees C and a \deltao O of $-1.7\pm 0.5\permil$. Our results, along with those from other clumped isotope studies of paleoenvironment, suggest that the \deltao O of seawater has remained broadly constant over time, in spite of a large potential for change through exchange of oxygen isotopes with crustal materials.

\section*{Acknowledgements}

We acknowledge Nami Kitchen, Kristin Bergmann, Daniel Stolper, and the rest of the Eiler lab for help with the clumped isotope measurements. We thank Chi Ma for his assistance with the SEM and the E-probe, and Joel Hurowitz for allowing us to use his ICP-AES. We also thank Mark Garcia and David Mann for their help with sample processing. This work was supported by NSF grant EAR-1053523 and the Agouron Institute.

\newpage

\section*{References}
\bibliographystyle{my-harv}
\bibliography{draft}

\begin{figure}[p]
\centering
\includegraphics[width=0.5\textwidth]{Fig_map.pdf}
\caption{Geologic map of the northern half of the island of Gotland, Sweden, modified from \cite{Calner2004a}. Approximate sample locations are marked by colored circles. Numbering corresponds to the sample location in Table \ref{results}.}
\label{map}
\end{figure}

\begin{figure}[htb]
\centering
\includegraphics[width=0.9\textwidth]{Fig_column-01.jpg}
\caption{On the left, a stratigraphic column showing the generalized lithstratigraphic units that were sampled for this study. On the right are the carbon and oxygen isotopic compositions of apparently well-preserved brachiopods, collected from the same section in Gotland. Isotopic compositions of brachiopods are modified from \cite{Bickert1997}.}
\label{column}
\end{figure}

\begin{sidewaysfigure}[htb]
\centering
\includegraphics[width=\textwidth]{Fig_section-01.jpg}
\caption{The \deltac C, \deltao O and $\Delta_{47}$ temperature of all the samples analyzed in this study, plotted against stratigraphic height. A height of zero meters is defined by the contact between the Lower Visby Beds and the Upper Visby Beds. Also included are data from Munnecke et al. (2003).}
\label{section}
\end{sidewaysfigure}

\begin{figure}[htb]
\centering
\includegraphics[width=\textwidth]{Fig_Brand-01.jpg}
\caption{Bulk manganese and strontium concentrations of the brachiopods, rugose corals, micrites and cements in this study. The brachiopods in this study have less Mn and more Sr than most of the brachiopods analyzed by Brand et al. (2012), which are also shown in this figure. Lines represent estimates by different authors of the range of pristine concentrations defined for brachiopods. Rugose corals generally have more Fe (not shown), more Mn and less Sr than brachiopods, giving them more altered trace metal signatures, while cements and micrites tend to have the most highly altered trace metal signatures.}
\label{Brand}
\end{figure}

\begin{figure}[htb]
\centering
\includegraphics[width=\textwidth]{Fig_metals-01.jpg}
\caption{Scatter plots showing the relationship between temperature and bulk trace metal concentration. The only statistically robust correlation ($\text{p}<0.05$ compared to zero correlation) is that between temperature and Sr concentration of rugose corals.}
\label{metals}
\end{figure}

\begin{figure}[htb]
\centering
\includegraphics[width=0.8\textwidth]{Fig_EBSD_coral-01.jpg}
\caption{Four different images of the same area of a rugose coral thin section, all at the same magnification. A) Reflected light. The line of dark pits are damaged caused by the E-probe electron beam. The dark triangle labelled ``RC'' was identified as a recrystallized domain of the coal septum. B) EBSD analysis. The recrystallized domain is clearly marked by many calcite crystals with the same orientation. The black scale bar represents 500 $\mu$m. C) Iron and D) manganese concentrations measured using the E-probe. The maps are qualitative, with warmer colors representing higher concentrations. The overlaid spots were measured quantitatively, and are labeled in units of ppm. For both metals, the recrystallized domain is indistinguishable from the rest of the septum.}
\label{EBSD_coral}
\end{figure}

\begin{figure}[htb]
\centering
\includegraphics[width=\textwidth]{Fig_EBSD_brach-01.jpg}
\caption{A) Transmitted polarized light image of a thin section of a brachiopod with $\Delta_{47}$ temperature 34\degrees C. The color overlay is an electron backscatter diffraction image of the same sample. The orientation of calcite crystals is indicated by color. B) Euler pole figure generated from the EBSD analysis.}
\label{EBSD_brach}
\end{figure}

\begin{figure}[htb]
\centering
\includegraphics[width=\textwidth]{Fig_flakes-01.jpg}
\caption{Scanning electron microscope images of flakes of brachiopod shell. Images were taken in secondary electron mode. Panels A--C show flakes that were judged to be pristine and are representative of the material used for analysis. Panel D shows a flake that was judged to be partially recrystallized and would not have been chosen for analysis. Clumped isotope temperatures of the samples for which representative flakes are shown are: A) 34 $\pm$ 2\degrees C; B) 35 $\pm$ 3\degrees C; C) 35 $\pm$ 3\degrees C.}
\label{flakes}
\end{figure}

\begin{figure}[htb]
\centering
\includegraphics[width=0.7\textwidth]{Fig_ellipses-01.jpg}
\caption{The $\Delta_{47}$ temperature and calculated water \deltao O for all samples analyzed in this study. Error ellipses represent one standard deviation of the mean of all measurement acquisitions for each sample.}
\label{ellipses}
\end{figure}

\begin{figure}[htb]
\centering
\includegraphics[width=0.5\textwidth]{Fig_comparisons-01.jpg}
\caption{Comparisons between the \deltac C, \deltao O and $\Delta_{47}$ temperature of brachiopod samples and adjacent micrite.}
\label{comparisons}
\end{figure}

\begin{figure}[htb]
\centering
\includegraphics[width=0.6\textwidth]{Fig_veizer-01.jpg}
\caption{In gray, the \deltao O$_{\text{PDB}}$ of apparently well-preserved skeletal calcites, compiled by \cite{Veizer1999}. In black are two different hypotheses for the trend in the \deltao O$_{\text{SMOW}}$ of the oceans through time.}
\label{Veizer}
\end{figure}

\end{document}